The Name of the Title is Hope

### 2023-12-13 input_marked_up/rebuttal/rebuttal_letter.tex

\section{comparison with baseline conditions}
% [1AC] The authors did a comparison between three conditions. However, how those conditions are compared to other techniques (as a baseline) in previous studies should be discussed.

% [2AC] The current baseline (EXTERNAL) condition is too weak. It is a trivial finding that the SoA of the AUGMENTED condition is higher than that of the EXTERNAL condition. Why has no comparison been made with a simpler technique where EMS is triggered by heuristic rules rather than based on EEG signals? Is there a reason the authors did not use as a baseline any technique suggested in previous studies?
% So it seems we should consider the INTENTION condition as a more important baseline for this study. And the fact that the AUGMENTED condition shows a lower SoA than the INTENTION condition (Figure 4) is a common problem with EMS-based techniques found in previous studies and is also a key motivation for this study. In other words, it appears that important problems have not been solved even though EEG signals have been utilized.
% How does the difference in SoA between the INTENTION condition and the AUGMENTED condition in previous studies compare to the difference in this study?

- [1AC, 2AC, R1]: comparison with other baseline conditions
The major motivation for the current work was to present a closed-loop augmentation system using neural input that is entirely decoupled from any outside constraints. Ultimately, we aim at designing a system that a user can wear and use in a variety of settings and applications as no knowledge about the user's environment should b enecessary for it to work. With this goal in mind, we did not find a useful baseline condition from previous studies to compare our interface to. We believe the simple case of random activation, for example 'send 10 EMS pulses within every 5 minutes' to not be a useful heuristic since we have shown that our classifier performs above chance level (F1 score of .7). Hence, as we have demonstrated, it outperforms activating EMS at chance level. Any other method to predict the user's intent to interact would be reliant on measuring some form of behavior or physiology, or constraining the interaction to a stimulus-response pattern.

% - Instead we chose to position the agency experience when using the prototype between the two edge cases of full control (INTENTION) vs. being passively moved (EXTERNAL).
% - because it would distract from the goal that this work was set out to tackle.
% - we want to have free behavior therefore there is no heuristic, or just a very basic one like activate 10 times every minute.
% 1. we focused on investigating sense of agency, not rt
% [ ] we adjusted the framing of the paper in the introduction accordingly and better situate it now within the action augmentation literature
% [ ] compare it in reference to earlier work and add a future outlook on subsequent validation steps against existing approaches and the potential fusion of these.

\section{on reduction in reaction time}

% [1AC] Both 2AC and R1 are concerned with a justification of the classification rate (10Hz) and the maximum interaction time reduction that this technique can achieve.

% [1AC] The used timing in the EXTERNAL condition was not particularly extreme and randomly picked from within 5-95\% of that user's previous timings. So they weren't faster here, and it's unclear whether they were faster in the augmented condition either.

% [2AC] I agree with the authors' idea that reaction time can be reduced through EMS if we know in advance when the user will move. However, according to the technique proposed by the authors, the maximum amount of reaction time we can reduce is the time interval between detection of the readiness potential (RP) and execution of the movement. Isn't this time interval usually very short (i.e., less than 200 ms)? Furthermore, for more robust classification, wouldn't EEG data sometimes need to be observed for a longer period of time, and as a result, the margin for reducing reaction time could be further reduced? We even have to consider the latency of the EMS itself.

% [2AC] Considering all these factors, I am curious to what maximum reaction time reduction the authors' proposed technique can actually achieve.

% [2AC] The time interval between the detection of RP and the actual movement of the participant's finger is very short (in Figure 5a), so isn't the classification rate of 10 Hz too low?

% [R1] But I'd like to start with something I found a bit odd. The paper positions itself within the human augmentation space, and I think that makes sense. The intro and method sections also discuss this. But did the augmented condition here actually improve reaction times? That does not actually become clear. As stated on p. 6, in the external condition, the used timing was not particularly extreme and randomly picked from within 5-95\% of that users previous timings. So they weren't actually faster here, and it's not clear whether they were faster in the augmented condition either. That might well not matter, given that this paper is not about reaction times per se. However, it is a curious aspect of the experimental design, reporting, and framing and thus would be good to discuss a bit more. Given that there probably would be a higher experienced loss of control with out of the ordinary reaction times, this might well have had an influence on the results, though.

- [1AC, 2AC, R1]: on the potential reduction in reaction time

The focus of the revised manuscript is now on SoA and is not deferred as much anymore by a secondary aspect of the system, i.e., a reduction in reaction times. In the revised manuscript we tightened the focus and framing on SoA and moved the part on previous studies reducing reaction times from the introduction section to related work. This better fits with the direct contrast of using predictive models (based on physiological signals) instead of controlled environments for action augmentation that falls in line with a user's intentions. Furthermore, we clarified reviewers' comments concerning the reduction in RT our system can achieve, i.e. prediction smoothing.

With the current implementation of the prototype, we saw an end-to-end latency from the measured physiological signals to the gating of the EMS switch of around 40 ms. Since this was implemented in relatively slow python code the time could be brought down significantly with more native code. While we could have applied the classifier at 20 Hz (one prediction every 50 ms) we wanted to be certain not to run into problems due to buffer overflow which is why we kept the real-time classification at 10 Hz. As 2AC points out, the RP precedes the movement onset by 200-300 ms. If the classifier works as intended, it should be able to pre-empt the movement, sending the first EMS pulse at 150 ms (200 - 50 ms) before the movement.

However, it is difficult to assess the reduction in reaction time at a high resolution as no label exists for the 'true' intention onset. This is the key contribution (and distinction) as compared to previous works in controlled environments. We are adding labels based on implicit classification of EEG signals in freely behaving users. Hence, we cannot assess the reduction based on any other metric other than the classifier, but if it works as designed, the reduction is there. However, we tried to explore the subjective experience of a potential reduction in RT through interviews. Another way would be to determine this on a single trial level by asking questions (explicit) or by inferring from neural signals (implicit). We added a paragraph to the discussion.

\section{technical details about the classifier}

By including a random effect for participants in our LME analyses, we accounted for the dependency of repeated measures within participants. The analysis is correct. We clarified [R2]'s comment about multiple comparison corrections and added details about why electrode Cz was used to report exemplary EEG results. Furthermore, we now report electrodes that were leveraged for classification in several participants and rephrased how we smoothed the prediction output [R1].

EMG and EMS were NOT recorded at the same time, as EMG was only recorded in the first condition to obtain labels for training data for the classifier. Following [2AC]'s suggestion, we added a new figure to reflect this and show the processing steps in detail. Additionally, we changed the order of the apparatus and procedure section, so readers have a better understanding of the setup before going into the task.

To improve readability and clarity [R2, R1] we now explain how the augmentation was achieved much earlier in our paper (both in the abstract and introduction). Furthermore, we better situate our augmentation system in the human augmentation/integration literature, thereby making our paper more accessible to a broader HCI audience. Lastly, we now include references to the distinction between pre-movement and idle EEG states.

% [R1] A big part of the work here is the classifier that determines whether a participant is idling or just about to move. I'm not an expert here, but I was left with a few questions regarding this classifier. For starters, it's not clear to me why the main reference throughout is Cz. The paper mentions this electrode earlier together with C3 and C4 as commonly used, but how does Cz stand out from the three? All the training is then in relation to the time of movement as detected by the EMG. So not when the brain started to plan the motion, but when the muscle actually is activating. As the whole second up to that moment is training data, it thus contains not just the planning either (but of course _also_ the planning). As Section 3.4.1 notes, 1s signal windows are condensed to slope values, based on the first and last 100ms. Per Section 3.4.3, this window is updated 10 times a second, so essentially shifted by 100ms. However, the predictions are smoothed with two previous ones (btw. I don't get the weighting: 0.3 & 0.5 for the previous predictions, and then presumably only 0.2 as weight for the current one? Why such a low value for the most up-to-date value?). However, doesn't that also mean that prediction of pre-movement is quite laggy? For a slope to be noticeable, the window must likely overlap quite a bit with the increase and then another 200ms of delay come to the averaging. With the EMS then active for 500ms, I'm pretty much wondering whether this did preempt the motion and whether the system then remained active longer than needed. As the paper itself states, the performance of the classifier is mixed with several participants reporting issues. So this is probably an aspect to provide a bit more details and explanations for.

% [ ] also why wasnt a link function used? assumed linear relationship between time estimation / questionnaires and study design

% [2AC] Most of the movements we perform in daily life are continuous rather than discrete, such as tapping. For example, human continuous movement (e.g., pointing) is known to be based on intermittent motor control [1]. In such cases, RP is expected to occur intermittently and repeatedly. Even in such cases, can the technique suggested by the authors be applied?
% -> unfold reference, % - unfold overlapping ERPs, for example with unfold toolbox

### 2023-12-13 input_marked_up/5_user_study_and_methods.tex
create duplicate marked up version

\section{User study \& Methods}
With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user's own brain signals serve as the control signal to a physical end effector. To this end, we assessed whether our prototype preserved agency using a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.
% todo add some more framing here to make it easier to step through the methods, maybe: For this paper we designed a prototype that lies between the two edge cases -- no intention and control to full control -- and investigate user's sense of agency using both explicit and implicit measures.

@@

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED.
% , see figure~\ref{fig:task_design}.

@@

\indent\textbf{INTENTION.} The task went as follows: (1) a fixation cross appeared on a tablet screen and participants were instructed to rest and wait until it disappeared; (2) then, they were instructed to wait for a brief moment (2 to 3s), before (3) initiating their movement and tap the screen. In line with the literature on the origin of the RP generating process they were told ``to avoid pre-planning the movement, avoid any obvious rhythm over trials, and to press when they felt the spontaneous urge to move''~\cite{Schultze-Kraft2021-cu}. (4) After the screen was tapped, a tone was played at a pseudo-random delay of 200, 350, or 500ms. Participants were now asked to estimate the delay, typing in their answer on a number pad of an attached keyboard. After confirming their answer by hitting the return key, the next trial started.
During INTENTION, participants were equipped with Electromyogram (EMG) sensors instead of EMS electrodes. INTENTION data was used as training data for the BCI, see section~\ref{BCI} below, as well as to select stereotypical reaction times in the EXTERNAL condition.

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both EXTERNAL and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview.

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel EMG device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected via two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses. To assist readers in replicating our experiment, we provide the necessary technical details, the complete source code for the experiment, the collected data, and the analysis scripts\footnote{[anonymized]}.
% TODO add links after anonymization is lifted.

\indent\textbf{(1) EMG Recording.} EMG data was recorded from 1 bipolar channel using the BrainAmp ExG amplifier (BrainProducts GmbH, Gilching, Germany). The two electrodes were placed above the flexor digitorum profundus with a reference electrode located on the wrist bone. EMG data was collected in synchrony with the EEG data through BrainProducts BrainVision Recorder.

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10$\Omega$. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/setup.png}
    \caption{Experimental setup of measurement-- and input devices (image with consent from participant).}
    \label{fig:setup}
\end{figure}

\indent\textbf{(3) Electrical Muscle Stimulation.} We actuated the ring finger via EMS, which was delivered with two electrodes attached to the participants' flexor digitorum profundus muscle. We utilized the flexor digitorum profundus since we found that we can robustly actuate it without inducing unintended motion of neighboring fingers. This finger actuation was achieved via a medically-compliant battery powered muscle stimulator (TENS/EMS Super Duo Plus, prorelax, Düren, Germany). The EMS system's output was controlled by flipping a solid state relay (silent) connected via an Arduino Uno (Arduino, Monza, Italy) to the experiment computer. The EMS was pre-calibrated by the participant to ensure a pain-free but effective stimulation and robust actuation leading to an `immediate' tap on the touchscreen after actuation. To ensure a comfortable experimental experience so that participants relaxed their arm musculature as much as possible, a custom built hand rest (support device) was placed on top of the touchscreen for participants, see figure~\ref{fig:setup}.
% [todo] What were the settings of the stimulation?

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an external keyboard to allow response input of users indicating their timing judgements.

% alternative headings:
% Brain-Computer Interface to Detect EEG Readiness Potentials }
% Classifiying EEG Signals of the Intent to Interact
\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. We utilized the [anonymized] and the EEGLAB~\cite{Delorme2004-sn} toolbox inside the MATLAB (The MathWorks Inc. Natick, MA, USA) environment for preprocessing both EEG and EMG data. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, in order to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset.
% bemobil pipeline

% TODO Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Preprocessing EEG and Selecting discriminative channels}\label{eeg_methods}
The EEG data was band pass filtered from 0.1 to 15 Hz. In the first step to prepare the EEG data for classification, noisy data segments were rejected. To this end, the EEGLAB function `autorej' was used, keeping default parameters. A trial was excluded if \textit{either} data from the \textit{idle} or \textit{pre-movement} class was rejected.

Then, to decrease the dimensionality of the EEG data, a selection of discriminative channels was conducted following a selection approach outlined by~\citep{Schultze-Kraft2021-cu}. First, for both \textit{pre-movement} and \textit{idle} the mean signal in the last 100 ms of each segment was subtracted from the mean signal in the first 100 ms of the segment. The resulting value thus indicated how much the signal had changed in the course of the 1 s segments. In line with the literature, there should be a change over time in the \textit{pre-movement} segment but not in the \textit{idle} segment. Then, all channels were sorted (1) in descending order by the signal difference in the \textit{pre-movement} segments and (2) in ascending order by the signal difference in the \textit{idle} segments. With respect to the ability of the system to discriminate between movement intention and the absence of movement intention, an ideal channel should show a strong change in signal over the course of a \textit{pre-movement} segment but no difference during an \textit{idle} segment. The ranks of the two criteria were joined by summation and the resulting order was saved for later use in training and real-time application of the classifier. Lastly, channels C3, C4 and Cz were moved to the top of the order for every participant as these are most frequently reported in studies on the RP.

\subsubsection{Training of EEG classifier}
To classify EEG data a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient feature of the selection of the most discriminative channels as features, hence in the end, generating a feature vector in the dimension of channels that were kept for classification. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting.

Using scikit-learn~\cite{Pedregosa2012-sj}, an LDA with automatic shrinkage was cross-validated (using 5-folds) for a grid search from 6 to 20 channels with a step size of adding 2 channels. Following this grid search, the cross-validation with the highest accuracy determined the number of channels that were kept for training the model for real-time application. Ultimately, to determine the threshold at which, during real-time application, the classifier would trigger the EMS, we computed the Receiver-operator characteristic (ROC) and from this selected the threshold at 15 \% false positive rate.

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed for the discriminative channels selected in the classifier cross-validation grid search. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output with the goal of reducing false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging the last two predictions with a weighting (.3 and .5). Then, at 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s.

@@

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}
To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. At the conclusion of each condition we prompted participants to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been we set the focus to the tapping movement and asked them to ignore the time estimation task for the following questions. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

% add back in after anonymization: All interviews were manually transcribed and then automatically translated using DeepL\footnote{https://www.deepl.com/translator}
We analyzed the interviews that always followed the AUGMENTED condition by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and [anonymized]. Prior to screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into positive and negative valence sentiments.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we report the F1 score, i.e. the harmonic mean between of the precision and recall, and plot the ROC.

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl}. Three time intervals were considered relevant to describe `regular' behavior across trials which would manifest by: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's methods to these criteria led to the exclusion of 102 trials across all participants (M = 34.00, SD = 16.46) and a final data set of 1698 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA. To test this, we fit a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect. As a random effect, we considered the intercepts for participants. We obtained p-values of the coefficients by calculating likelihood-ratio tests. All parameters were estimated by maximum likelihood estimation~\cite[see][]{Pinheiro2006-bk}. Post-hoc, pairwise contrasts were computed using estimated marginal means with Tukey's methods for multiple comparisons corrections~\cite{Lenth2020-xk}.
% ~\cite{winterLinearModelsLinear2013}

>>>>

\section{User study \& Methods}
\textcolor{green}{With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user's own brain signals serve as the control signal to a physical end effector. Therefore, we conducted a user study comparing three conditions of agency experience. With the first two conditions, INTENTION and EXTERNAL we queried users at the two edges of agency experience. In EXTERNAL participants had no intention to move and no control over their movement. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and EXTERNAL, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.}

\textcolor{red}{With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user’s own brain signals serve as the control signal to a physical end effector. To this end, we assessed whether our prototype preserved agency using a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.}

@@

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel EMG device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected to two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses. To assist readers in replicating our experiment, we provide the necessary technical details, the complete source code for the experiment, the collected data, and the analysis scripts\footnote{[anonymized]}.
% TODO add links after anonymization is lifted.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/setup_fixed.png}
    \caption{Experimental setup of measurement-- and input devices (image with consent from participant).}
    \label{fig:setup}
\end{figure}

\indent\textbf{(1) EMG Recording.} EMG data was recorded from 1 bipolar channel using the BrainAmp ExG amplifier (BrainProducts GmbH, Gilching, Germany). The two electrodes were placed above the flexor digitorum profundus with a reference electrode located on the wrist bone. EMG data was collected in synchrony with the EEG data through BrainProducts' BrainVision Recorder. \textcolor{green}{EMG data was only recorded for the first experimental condition to obtain labels for classifier training. After that, the EMG electrodes were changed for EMS electrodes.}

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes referenced to \textcolor{green}{electrode FCz} in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10k$\Omega$. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

\indent\textbf{(3) Electrical Muscle Stimulation.} We actuated the ring finger via EMS, which was delivered with two electrodes attached to the participants' flexor digitorum profundus muscle. We utilized the flexor digitorum profundus since we found that we can robustly actuate it without inducing unintended motion of neighboring fingers. This finger actuation was achieved via a medically compliant battery-powered muscle stimulator (TENS/EMS Super Duo Plus, prorelax, Düren, Germany). The EMS system's output was controlled by flipping a solid state relay (silent) connected via an Arduino Uno (Arduino, Monza, Italy) to the experiment computer. The EMS was pre-calibrated by the participant to ensure a pain-free but effective stimulation and robust actuation leading to an `immediate' tap on the touchscreen after actuation. To ensure a comfortable experimental experience so that participants relaxed their arm musculature as much as possible, a custom built hand rest (support device) was placed on top of the touchscreen for participants, see figure~\ref{fig:setup}.

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an external keyboard to allow \textcolor{red}{response input of users indicating}~\textcolor{green}{users to input} their timing judgements.

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED, see figure~\ref{fig:task_design}.

@@

\indent\textbf{INTENTION.} The task was as follows: (1) a fixation cross appeared on a tablet screen and participants were instructed to rest and wait until it disappeared; (2) they were instructed to wait for a brief moment (2 to 3s), before (3) initiating their movement and tap the screen, see figure~\ref{fig:progression}. In line with the literature on the origin of the RP generating process, they were told ``to avoid pre-planning the movement, avoid any obvious rhythm across trials, and to press when they felt the spontaneous urge to move''~\cite{Schultze-Kraft2021-cu}. (4) After the screen was tapped, a tone was played at a pseudo-random delay of 200, 350, or 500ms. Participants were now asked to estimate the delay, typing in their answers on a number pad of an attached keyboard. After confirming their answer by hitting the return key, the next trial started.

During INTENTION, participants were equipped with \textcolor{red}{Electromyogram (EMG)}\textcolor{green}{EMG} sensors instead of EMS electrodes. \textcolor{red}{The EEG data from INTENTION was used as training data for the BCI, see section~\ref{BCI} below}. Participants' reaction times in INTENTION were used to select stereotypical reaction times for the EXTERNAL condition. At the end of INTENTION, EMG electrodes were exchanged for EMS electrodes at the identical location on the forearm.

% \textcolor{green}{The EMG data was used to detect muscle activity onsets, i.e. movement onsets. The detected onsets were then used to label the data of the \textit{pre-movement} class, see section~\ref{BCI} below for more details.}

@@

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both EXTERNAL and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview~\textcolor{green}{immediately after the experience with the prototype.}

% alternative headings:
% Brain-Computer Interface to Detect EEG Readiness Potentials }
% Classifiying EEG Signals of the Intent to Interact
\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. We utilized the [anonymized] and the EEGLAB~\cite{Delorme2004-sn} toolbox inside the MATLAB (The MathWorks Inc. Natick, MA, USA) environment for preprocessing both EEG and EMG data. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset.
% TODO remove anonymization bemobil pipeline

% TODO/Ignore Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

@@

\subsubsection{Preprocessing EEG and Selecting discriminative channels}\label{eeg_methods}
The EEG data was band pass filtered from 0.1 to 15 Hz. In the first step to prepare the EEG data for \textcolor{red}{classification}~\textcolor{green}{classifier training}, noisy data segments were rejected. To this end, the EEGLAB function `autorej' was used, keeping default parameters. A trial was excluded if \textit{either} data from the \textit{idle} or \textit{pre-movement} class was rejected.

Then, to decrease the dimensionality of the EEG data, we selected discriminative channels following an approach outlined by~\citep{Schultze-Kraft2021-cu}. First, for both \textit{pre-movement} and \textit{idle} the mean signal in the last 100 ms of each segment was subtracted from the mean signal in the first 100 ms of the segment. The resulting value thus indicated how much the signal had changed in the course of the 1 s segments. In line with the literature, there should be a change over time in the \textit{pre-movement} segment but not in the \textit{idle} segment. Then, all channels were sorted (1) in descending order by the signal difference in the \textit{pre-movement} segments and (2) in ascending order by the signal difference in the \textit{idle} segments. With respect to the ability of the system to discriminate between movement intention and the absence of movement intention, an ideal channel should show a strong change in signal over the course of a \textit{pre-movement} segment but no difference during an \textit{idle} segment. The ranks of the two criteria were joined by summation and the resulting order was saved for later use in training and real-time application of the classifier. Lastly, channels C3, C4, and Cz were moved to the top of the order for every participant as these are most frequently reported in studies on the RP.

\subsubsection{Training of EEG classifier}
To classify EEG data, a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient of the selection of the most discriminative channels as features, generating a feature vector in the dimension of channels that were kept for classification. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting~\textcolor{green}{and decrease the computational load.}

Using scikit-learn~\cite{Pedregosa2012-sj}, the LDA \textcolor{red}{with automatic shrinkage} was cross-validated (using 5-folds) for a grid search from 6 to 20 channels with a step size of adding 2 channels. Following this grid search, the cross-validation with the highest accuracy determined the number of channels that were kept for training the model for real-time application. Ultimately, to determine the threshold at which, during the real-time application, the classifier would trigger the EMS, we computed the Receiver-operator characteristic (ROC) and from this selected the threshold at 15 \% false positive rate.

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected \textcolor{green}{discriminative channels}. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed. \textcolor{red}{for the discriminative channels selected in the classifier cross-validation grid search.} This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output to reduce false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging the \textcolor{red}{last two} \textcolor{green}{current and the preceding prediction with a weighting (weight of .3 for the preceding, and .5 for the current prediction). The weights were obtained through trial-and-error during piloting}. Then, with a 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s.

@@

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}
To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. After each condition we prompted participants to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been, we set the focus to the tapping movement and asked them to ignore the time estimation task in the questions that followed. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

% add back in after anonymization: All interviews were manually transcribed and then automatically translated using }
We analyzed the interviews \textcolor{red}{that always followed the AUGMENTED condition} by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and translated to English using DeepL\footnote{https://www.deepl.com/translator} (DeepL SE, Cologne, Germany). Before screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants' estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into sentiments with positive and negative valence.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. \textcolor{green}{Electrode Cz was chosen for exemplary presentation, as it is frequently reported in studies on the RP, since it captures neural activity originating from the sensorimotor cortex.} Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we calculated the F1 score, i.e. the harmonic mean of precision and recall, and plotted the ROC\textcolor{green}{, see figure~\ref{fig:EEG_results} a to c.}

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl}. Three time intervals were considered relevant to describe `regular' behavior across trials which would manifest by: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's method \textcolor{green}{to each of these time windows and fusing the rejected trials} \textcolor{red}{criteria} led to the exclusion of 102 trials across all participants (M = 34.00, SD = 16.46) and a final data set of 1698 trials.
